---
title: "Whack-a-Mole Eye Tracking Analysis"
author: "Bastian Ilsø Hougaard, Hendrik Knoche"
date: "3/27/2020"
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
options(scipen=1, digits=2)
dir = 'csv_data'

D_game = read.csv(file = '../csv_data/gameLog.csv')
D_gaze = read.csv(file = '../csv_data/gaze_positions.csv')
D_pupil = read.csv(file = '../csv_data/pupil_positions.csv')
D_fixations = read.csv(file = '../csv_data/fixations.csv')

D_all = D_game %>% full_join(D_gaze, by = c("PupilTime" = "gaze_timestamp")) %>% arrange(PupilTime)

```

## Table of Contents
1. [Introduction](#introduction)
3. [Oculometrics](#oculometrics)

## Introduction

This document provides analysis of data collected by Whack-A-Mole VR and Pupil Capture:

 * `gameLog.csv` is Whack-A-Mole's game data.
 * `export_info.csv` is generic export information from Pupil Capture.
 * `fixation_report.csv` is Pupil Player's aggregated summary of counted fixations.
 * `fixations.csv` contains every fixation found during the game.
 * `pupil_positions.csv` contains every frame of data from the eye tracking cameras.
 * `gaze_positions.csv` contains Pupil's interpreted 3D gaze data (how it maps to VR).
 * `world_timestamps.csv` contains mapping from pupil_time to frame in the recorded video.
 * `world.mp4` is the annotated video.


## Oculometrics

### TODO

 * Calculate things with `PupilSize` - maybe for Pupil Dilation..
 * Saccade analysis (?)

### Summarizing Eye movements as visual angles.

In eye tracking analysis we prefer to analyze data using established units such as visual angles.

#### Using PFA's suggestion

The below results is the based on using PFA's calculation method. His calculation method assumes that the origin position of the eye is `0,0,0`. Therefore distance is simply assumed to be the vectors' magnitude.

**Math:** `angle = arccos( (a dot b) / (|a| * |b|) )`


```{r visual-angle-1, echo=FALSE}
visAngle <- function(a_x, a_y, a_z, b_x, b_y, b_z){
  # this function calculates visual angle
  # size and distance must be in the same units
  a = c(a_x, a_y, a_z)
  b = c(b_x, b_y, b_z)
  if (anyNA(a) || anyNA(b)) {
    return(NA)
  }
  
  Rad = acos( (a %*% b) / ( norm(a, type="2") * norm(b, type="2") ) )
  Ang = Rad*(180/pi)
  return(Ang)
}

D_gaze$vis_degrees = mapply(visAngle,lag(D_gaze$gaze_point_3d_x), lag(D_gaze$gaze_point_3d_y), lag(D_gaze$gaze_point_3d_z), D_gaze$gaze_point_3d_x, D_gaze$gaze_point_3d_y, D_gaze$gaze_point_3d_z)


summaries = D_gaze %>%
            summarize(min = min(vis_degrees, na.rm=T),
                      quantile1 = quantile(vis_degrees, 0.25, na.rm=T),
                      mean = mean(vis_degrees, na.rm=T),
                      quantile3 = quantile(vis_degrees, 0.75, na.rm=T),
                      max = max(vis_degrees, na.rm=T))

kable(summaries) # formats the output as a nice table.

example_coord =  data.frame(
  point_a = paste(round(D_gaze$gaze_point_3d_x[1],2),round(D_gaze$gaze_point_3d_y[1],2),round(D_gaze$gaze_point_3d_z[1],2), sep=", "),
  point_b = paste(round(D_gaze$gaze_point_3d_x[2],2),round(D_gaze$gaze_point_3d_y[2],2),round(D_gaze$gaze_point_3d_z[2],2), sep=", "),
  angle = visAngle(D_gaze$gaze_point_3d_x[1], D_gaze$gaze_point_3d_y[1], D_gaze$gaze_point_3d_z[1], D_gaze$gaze_point_3d_x[2], D_gaze$gaze_point_3d_y[2], D_gaze$gaze_point_3d_z[2])
  )

kable(example_coord)

hist(D_gaze$vis_degrees)

```


#### Using Stephen Rhode's Implementation

Stephen Rhode's implementation uses `size` and `distance` to calculate visual angle. In this case, we determine `size` as magnitude of gaze position `gaze_point_3d` from previous gaze position, and `distance` as magnitude of gaze position from eye position `eye_center0_3d`.

**Math:** `angle = 2*arctan( size / (2 * distance) )`


```{r visual-angle-stephen, echo=FALSE}
calcSize <- function(a_x, a_y, a_z, b_x, b_y, b_z) {
  a = c(a_x, a_y, a_z)
  b = c(b_x, b_y, b_z)
  if (anyNA(a) || anyNA(b)) {
    return(NA)
  }  
  size = norm(b - a, type="2")
  return(size)
}

calcDistance <- function(b_x, b_y, b_z, e_x, e_y, e_z) {
  e = c(e_x, e_y, e_z)
  b = c(b_x, b_y, b_z)
  if (anyNA(b) || anyNA(e)) {
    return(NA)
  }  
  distance = norm(b - e, type="2")
  return(distance)
}

visAngle <- function(a_x, a_y, a_z, b_x, b_y, b_z, e_x, e_y, e_z){
  a = c(a_x, a_y, a_z)
  b = c(b_x, b_y, b_z)
  e = c(e_x, e_y, e_z)
  if (anyNA(a) || anyNA(b) || anyNA(e)) {
    return(NA)
  }  
  
  size = calcSize(a_x, a_y, a_z, b_x, b_y, b_z) #norm(b - a, type="2")
  distance = calcDistance(b_x, b_y, b_z, e_x, e_y, e_z) #(b - e, type="2")
  
  # this function calculates visual angle
  # size and distance must be in the same units
  Rad = 2*atan(size/(2*distance))
  Ang = Rad*(180/pi)
  return(Ang)
}

D_gaze$vis_eye1 = mapply(visAngle,lag(D_gaze$gaze_point_3d_x), lag(D_gaze$gaze_point_3d_y), lag(D_gaze$gaze_point_3d_z), D_gaze$gaze_point_3d_x, D_gaze$gaze_point_3d_y, D_gaze$gaze_point_3d_z, D_gaze$eye_center0_3d_x, D_gaze$eye_center0_3d_y, D_gaze$eye_center0_3d_z)
D_gaze$vis_eye2 = mapply(visAngle,lag(D_gaze$gaze_point_3d_x), lag(D_gaze$gaze_point_3d_y), lag(D_gaze$gaze_point_3d_z), D_gaze$gaze_point_3d_x, D_gaze$gaze_point_3d_y, D_gaze$gaze_point_3d_z, D_gaze$eye_center1_3d_x, D_gaze$eye_center1_3d_y, D_gaze$eye_center1_3d_z)


summary1 = D_gaze %>%
            summarize(eye1_min = min(vis_eye1, na.rm=T),
                      eye1_quantile1 = quantile(vis_eye1, 0.25, na.rm=T),
                      eye1_mean = mean(vis_eye1, na.rm=T),
                      eye1_quantile3 = quantile(vis_eye1, 0.75, na.rm=T),
                      eye1_max = max(vis_eye1, na.rm=T))

kable(summary1) # formats the output as a nice table.

summary2 = D_gaze %>%
            summarize(eye2_min = min(vis_eye2, na.rm=T),
                      eye2_quantile1 = quantile(vis_eye2, 0.25, na.rm=T),
                      eye2_mean = mean(vis_eye1, na.rm=T),
                      eye2_quantile3 = quantile(vis_eye2, 0.75, na.rm=T),
                      eye2_max = max(vis_eye2, na.rm=T))

kable(summary2) # formats the output as a nice table.

example_coord =  data.frame(
  point_a = paste(round(D_gaze$gaze_point_3d_x[1],2),round(D_gaze$gaze_point_3d_y[1],2),round(D_gaze$gaze_point_3d_z[1],2), sep=", "),
  point_b = paste(round(D_gaze$gaze_point_3d_x[2],2),round(D_gaze$gaze_point_3d_y[2],2),round(D_gaze$gaze_point_3d_z[2],2), sep=", "),
  eye1 = paste(round(D_gaze$eye_center0_3d_x[2],2),round(D_gaze$eye_center0_3d_y[2],2),round(D_gaze$eye_center0_3d_z[2],2), sep=", "),
  size = calcSize(D_gaze$gaze_point_3d_x[1], D_gaze$gaze_point_3d_y[1], D_gaze$gaze_point_3d_z[1], D_gaze$gaze_point_3d_x[2], D_gaze$gaze_point_3d_y[2], D_gaze$gaze_point_3d_z[2]),
  distance = calcDistance(D_gaze$gaze_point_3d_x[2], D_gaze$gaze_point_3d_y[2], D_gaze$gaze_point_3d_z[2], D_gaze$eye_center0_3d_x[2], D_gaze$eye_center0_3d_y[2], D_gaze$eye_center0_3d_z[2]),
  angle = visAngle(D_gaze$gaze_point_3d_x[1], D_gaze$gaze_point_3d_y[1], D_gaze$gaze_point_3d_z[1], D_gaze$gaze_point_3d_x[2], D_gaze$gaze_point_3d_y[2], D_gaze$gaze_point_3d_z[2], D_gaze$eye_center0_3d_x[2], D_gaze$eye_center0_3d_y[2], D_gaze$eye_center0_3d_z[2])
  )

kable(example_coord)

hist(D_gaze$vis_eye1)
hist(D_gaze$vis_eye2)

```

